Manuscript: Detecting differences in Size Spectra

### 2023-02-23 Manuscript.tex
MS draft 23 feb 2023

\title{Manuscript: Detecting differences in Size Spectra}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Vojsava
Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
\date{07 January, 2023}

@@

\textbf{NOTE} I'm still working on integrating with Zotero and importing
all the necessary references (thanks Jim for the help and
instructions!). Don't worry about the \textbf{??} references in the
rendered version for now.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The distribution of body size in communities is remarkably consistent
  across habitats and taxa and can be represented by size spectra
  relationships.
\item
  Many methods have been proposed for constructing size spectra
  relationships, but recent work has shown that varied methods return
  biased estimates of relationship parameters, and different methods in
  fact are not estimating the same parameter. Despite this variability
  in estimates, it is unclear if the relative change across
  environmental gradients is consistent across methodologies. Here, we
  simulate data sets across an hypothetical environmental gradient and
  estimate the size spectra parameter (slope or exponent, depending on
  method) at each site, and, importantly, estimate the relationship of
  parameters across the hypothetical gradient. We also use two
  previously published body size datasets across an anthropogenic stress
  gradient and an environmental temperature gradient and assess how the
  conclusions of these studies would vary based on methods used.
\item
  We find that maximum likelihood methods always perform better than
  common binning methods. Additionally, the variance in estimates using
  MLE methods is markedly reduced when compared to binning methods.
\item
  The uncertainty and variation in estimates when using binning methods
  is often greater than or equal to the variation previously published
  in experimental and observational studies, bringing into question the
  effect size of previously published results. However, re-analysis of
  two previously published datasets does not markedly alter the
  conclusions reached. Further study is needed to identify how and when
  these estimates may be biased, and when the general pattern of results
  can be considered to be ``true.''
\end{enumerate}

@@

Body size distributions are a fundamental characteristic of communities.
In general, size-abundance relationships are expected to scale with the
value of -0.75 across all biological communities (Damuth 1981, 1991,
1998). This is thought to be a consequence of simple size-dependent
metabolic constraints on organisms' energy use predicted by the
metabolic theory of ecology (Nee 1991, Brown et al.~2004). The
remarkable consistency of these relationships across spatiotemporal
scales and ecosystems has led them to be recommended as a ``universal''
indicator of ecological status (Petchey and Belgrano 2010)(Petchey and
Belgrano 2010). Variation in size-abundance relationships have been
documented through space (Pomeranz 2021), time (newish Sprules papers?),
seasonality (McGarvey and Kirk), to human activities (fish abundance
with no fishing paper). Likwise, variation in size spectra relationships
have been used to explain fundamental differences in how communities are
organized. For example, external resource subsidies ``bend the rules''
and allow higher abundances of large body sizes than would be expected
(Perkins). However, recent research has shown that these results may be
an artifact of how the data were treated. Edwards et al.~(2020?)
reanalyzed a time series of marine fisheries data and found that
previously reported changes through time were actually dependent on the
methodolugy used.

Individual size distributions (ISD \emph{sensu} (White et al. 2007)),
also referred to as abundance size spectra) are one of the
size-abundance relationships commonly used. ISDs show the relationship
between the abundances of individual body size regardless of taxonomy.
Generally, there is a negative relationship between individual body size
(M) on the \emph{x}-axis and abundance (N) on the \emph{y}-axis.
Theoretical and empirical data support this relationship being described
as a simple power law with exponent \(\lambda\) in the form of
\(N \sim M^{\lambda}\), where \(\lambda = 2\). Commonly, \(N\) is the
count of body sizes grouped into bins, and \(\lambda\) is estimated as
the slope from OLS regressions of log transformed data of \(N\) and the
mid point of the body size bin \(M_{bin}\):
\(\log(N_{count}) = \lambda \log(M_{bin})\). Myriad binning methods have
been proposed, including linear and logarithmic bin widths. Likewise,
methods can rely on the absolute count in the bins as well as
normalization techniques where the count is divided by the bin width
(especially common with logarithmic binning). Alternatively, \(\lambda\)
can be estimated directly using maximum likelihood techniques.

Previous work has shown that the estimates of \(\lambda\) differ between
MLE and OLS techniques. OLS methods are particularly sensitive to
decisions made in the binning process. Simulation studies have shown
that MLE offers consistently more accurate estimates of \(\lambda\)
(White et al.~2008, Edwards et al.~2017), and reanalysis of empirical
data also indicates that the conclusions are dependent on the
methodology used Edwards et al. (2020). However, recent empirical
analysis of stream macroinvertebrate communities across the National
Ecological Observatory Network (NEON, USA) showed that while the
estimates of \(\lambda\) varied, the relative change across the
environmental gradient was consistent regardless of method used
(Pomeranz et al. 2022). While there is a growing consensus that MLE
methods offer more reliable estimates of \(\lambda\), and binning
methods result in biased estimates, it remains unclear if these biases
are consistent and systematic or stochastic, and whether or not the
relative change in ISD parameters is consistent across space and time.
In other words, if the data within a study are all treated the same,
does a relative change of OLS slope parameters of 0.1 coincide with a
relative change of MLE estimates of 0.1?

We had three primary objectives in this study: 1) how well do different
methods estimate site-specific \(\lambda\)'s? 2) Do the estimates of the
hypothetical relationship obtained from different methods agree? and 3)
Are the conclusions reached from previously published data sets
dependent on the method used?

In order to answer the first two questions, we simulate body size
observations from bounded power law distributions with varied
\(\lambda\) exponents across a hypothetical environmental gradient and
compare the results obtained using three different methodologies common
in the literature. To answer the third question, we re-analyze two
previously published data sets of stream community body sizes across a
stress and environmental gradient, respectively. We find that the MLE
method more accurately estimates the site-specific \(\lambda\) exponents
as well as the relative change in exponents across the hypothetical
gradient, as well as having smaller variation around the estimates. The
logarithmic binning methods generally perform well, but have larger
variation around the site-specific estimates. In addition, the variation
of the estimates of the hypothetical relationship obtained from the
binning methods is dependent on the value of \(\lambda\) used (under and
over estimates steep and shallow \(\lambda\), respectively).

@@

In order to investigate the performance of commonly used methods, we
simulate body size observations from a bounded power law distribution
using the inverse method, as described in (Edwards et al. 2017). Let
\(M\) be a random variable of body sizes described by the probability
density function:

@@

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
on empirical body sizes of stream benthic communities reported in
(Pomeranz et al. 2022). Our results are not dependent on the range of
body sizes, and we show the results of other body size ranges in the
supplemental information.

@@

We sampled \(n = 1000\) body sizes from distributions described by five
different \(\lambda\)'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The values
of \(\lambda\) describe how quickly the abundance of large body sizes
decline within a community, e.g., a value of -1.5 means there would be a
relatively high number of large body sizes (shallow) whereas a value of
-2.5 means there would be relatively very few large body sizes (steep).
We used three different methods (described below) to estimate the value
of \(\lambda\). We repeated this process 1000 times (reps) and plot the
distribution of values obtained for each method.

@@

For each sample of \(M\) and each of the 1000 replicates, we estimated
the exponent \(\lambda\) directly using MLE methods modified from the
the \texttt{sizeSpectra} package (Edwards et al.~2020). Throughout the
manuscript, these estimates are referred to as MLE.

@@

In addition, we use two common binning methods to estimate the OLS slope
parameter in log-log space. It is important to note that the binning
methods are estimating a slope (\(\beta_{1}\)) from an OLS regression
analysis, and not \(\lambda\) directly. However, in the interest of
clarity, we will refer to all estimates as \(\lambda\) regardless of the
method used, and reserve the term \(\beta_{env}\) to refer to the
estimate of the size spectra relationships across the hypothetical
environmental gradient.

For the first binning method, we created 6 equal logarithmic bins
covering the range of body sizes. This method has been used extensively
in previous studies. For example see Dossena et al. (2012). In the
present study, the count in each bin was normalized by dividing by the
bin width to account for the unequal bin sizes. The process of
normalizing shifts the estimate by -1. In other words, an un-normalized
estimate of -0.75 would result in an estimate of -1.75 when normalizing
the data (Sprules and Barth 2015 pp. Edwards2018, Pomeranz2022).
Throughout the manuscript, the normalized equal logarithmic binning
method will be referred to as ELBn.

@@

The second method was similar to ELBn, but bins of Log\textsubscript{2}
widths are used. The number of bins is dependent on the size ranges
present in the data. When working with empirical data with different
size ranges this can alter the number of bins per site. However, since
the data here were simulated from a known size range, the number of bins
for each site is identical. The count in each bin is normalized in the
same way. This method has been used {[}Jennings and Blanchard (2004),
Jennings et al. (2002), Sprules and Barth (2015), Pomeranz et al.
(2019a); McGarvey and Kirk (2018)) and is referred to as the Normalized
Abundance Spectrum (NAS).

For the two binning methods, simple OLS regression were conducted in the
form \(Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)\). Where
\(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).

@@

\hypertarget{hypothetical-relationshiop-of--0.5}{%
\subsubsection{Hypothetical relationshiop of
-0.5}\label{hypothetical-relationshiop-of--0.5}}

@@

To do this, we again sampled body sizes from distributions described by
five different \(\lambda\)'s. However, because the estimates from the
binning methods varied depending on the value of \(\lambda\) (see
results), we chose three different ``windows'' of lambda values. The
three windows were steep
(\(\lambda = [-1.50,~ -1.75,~ -2.00,~ -2.25, -2.50]\)), medium
(\(\lambda = [-1.00,~ -1.25,~ -1.50,~ -1.75,~ -2.00]\)), and shallow
((\(\lambda = [-0.50,~ -0.75,~ -1.00,~ -1.25,~ -1.50]\)).

\textbf{NOTE} This MS seems to be getting long. Possibly present
``medium'' window in main results, and include other windows in SI

Each sample of body sizes from different \(\lambda\) values was assumed
to come from a community across a hypothetical environmental gradient
\(X\). Values of \(X\) for each community are uniformly distributed from
-1 to 1 (site-specific x-values: \((-1.0, -0.5, 0, 0.5, 1.0)\). For each
\(\lambda\) ``window,'' the relative change in \(\lambda\) values
(\(\beta_{env}\)) across the gradient is equal to -0.5, such that our
model is:

\textbf{JEFF} Is this correct? for some reason this seems wrong to me.

\[\beta_{env} = Norm(\alpha -0.5X,~\sigma)\]

We repeated the data simulation 1000 times (reps). The main results
presented here were not dependent on the range of \emph{x}-values or the
number of sites (supplemental information).

\hypertarget{varying-hypothetical-relationship}{%
\subsubsection{Varying Hypothetical
Relationship}\label{varying-hypothetical-relationship}}

Finally, we also varied the relationship (\(\beta_{env}\)) across the
hypothetical gradient. We tested \(\beta_{env}\) values of 0, -0.25, and
-0.5 by sampling 5 body sizes communities uniformly distributed across
the following sets of \(\lambda\) values: {[}-1.5{]}, {[}-1.25,
-1.75{]}, {[}-1, -2{]}.

\hypertarget{estimation-of-relationship-across-gradient-beta_env}{%
\subsection{\texorpdfstring{Estimation of relationship across gradient,
\(\beta_{env}\)}{Estimation of relationship across gradient, \textbackslash beta\_\{env\}}}\label{estimation-of-relationship-across-gradient-beta_env}}

For each simulation replicate and size spectra method, we estimated how
the size spectra parameters varied across the hypothetical gradient. We
did this by performing simple OLS regression of the
\(\lambda_{estimate}\) across the hypothetical gradient (\(X\)):

\[\lambda_{estimate} = \beta_0 + \beta_{env} * X\]

The distribution of the relationship coefficient, \(\beta_{env}\), were
plotted compared to the known relationship.

@@

We re-analyze two data sets of benthic macroinvertebrate communities
from stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage (AMD) stress gradient. Details of the
sample collection and processing can be found in (Pomeranz et al.
2019a). Briefly, all individuals from each sample were identified to the
lowest practical taxonomic unit and body lengths were measured using
image processing software from photos taken with a camera mounted to a
dissecting microscope. Body mass was estimated using taxon-specific
published length weight regressions.

The second data set was from the wadeable stream sites of the National
Ecological Observatory Network (National Ecological Observatory Network
(NEON) 2022). NEON stream sites are located across a wide temperature
gradient in the United States, from Puerto Rico to Alaska. Quantitative
macroinvertebrate samples were collected using the most appropriate
method based on the local habitat. All individuals were identified and
had their body lengths measured, and body mass was estimated using
published length weight regressions. This data has been analyzed
previously using size spectra methods as described in (Pomeranz et al.
2022). Detailed methods of the sampling collection and data processing
methods can be found on the NEON website
({[}\url{https://www.neonscience.org/data-collection/macroinvertebrates}{]},
or macroinvertebrate DPI pubs). Coefficient estimates (\(\pm\) 1
standard deviation) are compared across methods. This allows us to
determine whether or not the main results would differ depending on the
method used.

@@

The data used in this manuscript is already publicly available (Pomeranz
et al. 2019b) data dryad DOI
\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s},
(National Ecological Observatory Network (NEON) 2022). R Scripts to
reproduce the full simulation and analysis are available at:
({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).

@@

There was considerable variation in the \(\lambda\) estimate across
methodologies (Figure 1). The distribution of estimates from the MLE
method was almost always symmetrical and centered at the known value of
\(\lambda\) (Figure 1, blue distribution and dashed line). The
distribution of estimates from the binning methods were generally wider,
and sometimes showed assymetrical distributions.

@@

@@

The distribution of relationship estimates from the MLE method (Figure
3, blue) is approximately centered at the known relationship value in
each of the ``windows.'' The distribution for the MLE estimate is widest
in the ``steep'' window, but is still markedly narrower than the binning
methods. In addition to having much wider distributions, the binning
methods systematically overestimate the known relationship.

@@

Finally, we were interested in finding if the three methods performed
differently with varying values of the known relationship. Here, we
present estimates from each method when the known relationship
(\(\beta_{env}\)) across the hypothetical gradient was -0.5, -0.25, and
0, respectively (Figure 4).

\begin{figure}
\centering
\includegraphics{figures/vary_beta_plot.png}
\caption{Figure 4. Individual regressions for each rep (N = 1000) and
method (columns) for three different known relationship values (rows):
0, 0.25, and 0.5}
\end{figure}

>>>>

\title{Manuscript: Detecting differences in Size Spectra}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Vojsava
Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
\date{23 February, 2023}

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The distribution of body size in communities is remarkably consistent
  across habitats and taxa and can be represented by size spectra, which
  are described by a power law. The focus of size spectra analysis is to
  estimate the exponent (\(\lambda\)) of the power law.
\item
  Many methods have been proposed for estimating \(\lambda\) most of
  which involve binning the data, summing abundance within bins, and
  then fitting a ordinary least squares (OLS) regression in log-log
  space. However, recent work has shown that binning procedures may
  return biased estimates of size spectra exponents compared to
  procedures that directly estimate \(\lambda\) using maximum likelihood
  estimation (MLE). Despite this variability in estimates, it is unclear
  if the relative change across environmental gradients is consistent
  across methodologies. Here, we used simulation to compare the ability
  of two binning methods (NAS, ELBn) and MLE to 1) recapture known
  values of \(\lambda\), and 2) recapture parameters in a linear
  regression measuring the change in \(\lambda\) across a hypothetical
  environmental gradient. We also compared the methods using two
  previously published body size datasets across a pollution gradient
  and a temperature gradient
\item
  Maximum likelihood methods always performed better than common binning
  methods, which demonstrated consistent bias depending on the simulated
  values of \(\lambda\). This bias carried over to the regressions,
  which were more accurate when \(\lambda\) was estimated using MLE
  compared to the binning procedures. Additionally, the variance in
  estimates using MLE methods is markedly reduced when compared to
  binning methods.
\item
  The uncertainty and variation in estimates when using binning methods
  is often greater than or equal to the variation previously published
  in experimental and observational studies, bringing into question the
  effect size of previously published results. However, while the
  methods produced different slope estimates from previously published
  datasets, they were in qualitative agreement on the sign of those
  slopes (i.e., all negative or all positive). Our results provide
  further support for the direct estimation of \(\lambda\) using MLE (or
  similar procedures) over the more common methods of binning.
\end{enumerate}

@@

Body size distributions are a fundamental characteristic of communities.
In general, size-abundance relationships are expected to scale with the
value of -0.75 across all biological communities (Damuth 1981, 1991,
Damuth 1998). This is thought to be a consequence of simple
size-dependent metabolic constraints on organisms' energy use predicted
by the metabolic theory of ecology (Nee et al. 1991, Brown et al. 2004).
The remarkable consistency of these relationships across spatiotemporal
scales and ecosystems has led them to be recommended as a ``universal''
indicator of ecological status (Petchey and Belgrano 2010). Variation in
size-abundance relationships have been documented through space
(Pomeranz et al. 2022), time (Evans et al. 2022), seasonality (McGarvey
and Kirk 2018), and in response to human activities (Jennings and
Blanchard 2004). Likewise, variation in size spectra relationships have
been used to explain fundamental differences in how communities are
organized. For example, external resource subsidies ``bend the rules''
and allow higher abundances of large body sizes than would be expected
(Perkins et al. 2018). However, recent research has shown that these
results may be an artifact of how the data were treated. Edwards et
al.~(2020) reanalyzed a time series of marine fisheries data and found
that previously reported changes through time were actually dependent on
the methodology used.

Individual size distributions (ISD \emph{sensu} (White et al. 2007),
also referred to as abundance size spectra) are one of the
size-abundance relationships commonly used. ISDs represent frequency
distributions of individual body sizes within a community, regardless of
taxonomy. Generally, there is a negative relationship between individual
body size (M) on the \emph{x}-axis and abundance (N) on the
\emph{y}-axis. Theoretical and empirical data support this relationship
being described as a simple power law with exponent \(\lambda\) in the
form of \(N \sim M^{\lambda}\), where \(\lambda = 2\) (Sheldon and Kerr
1972, Andersen and Beyer 2006). Commonly, \(N\) is the count of body
sizes grouped into bins, and \(\lambda\) is estimated as the slope from
OLS regressions in log-log space of \(N\) and the mid point of the body
size bin \(M_{bin}\): \(\log(N_{count}) = \lambda \log(M_{bin})\).
Myriad binning methods have been proposed, including linear and
logarithmic bin widths. Likewise, some methods rely on the absolute
counts in the bins and others employ normalization techniques such as
dividing the count by the bin width (especially common with logarithmic
binning). As an alternative to binning methods, \(\lambda\) can be
estimated directly on un-binned data using maximum likelihood
techniques.

Previous work has shown that the estimates of \(\lambda\) differ between
MLE and size-binned OLS techniques. Size-binned OLS methods are
particularly sensitive to decisions made in the binning process.
Simulation studies have shown that MLE offers consistently more accurate
estimates of \(\lambda\) (White et al. 2008, Edwards et al. 2017), and
reanalysis of empirical data also indicates that the conclusions are
dependent on the methodology used (White et al. 2008, Edwards et al.
2020). However, recent empirical analysis of stream macroinvertebrate
communities across the National Ecological Observatory Network (NEON,
USA) showed that while the estimates of \(\lambda\) varied based on
method, the relative change across the environmental gradient was
consistent across methods (Pomeranz et al. 2022). While there is a
growing consensus that MLE methods offer more reliable estimates of
\(\lambda\), and binning methods result in biased estimates, it remains
unclear if these biases are consistent and systematic or stochastic, and
whether or not the relative change in ISD parameters is consistent
across space and time. In other words, if the data within a study are
all treated the same, does a relative change of size-binned OLS slope
parameters of 0.1 coincide with a relative change of MLE \(\lambda\)
estimates of 0.1?

We had Two primary objectives in this study: 1) to compare how well
different methods estimate site-specific \(\lambda\)'s 2) To compare how
different methods of estimating \(\lambda\) impact subsequent
regressions that use \(\lambda\)'s as a response variable. We did this
using both simulated and empirical data sets.

@@

To investigate the performance of commonly used methods, we simulate
body size observations from a bounded power law distribution using the
inverse method, as described in (Edwards et al. 2017). Let \(M\) be a
random variable of body sizes described by the probability density
function:

@@

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values were
based on empirical body sizes of stream benthic communities reported in
(Pomeranz et al. 2022). Our results were not dependent on the range of
body sizes (Supplemental Information).

@@

Using the procedure above, we simulated \(n = 1000\) body sizes from
five different \(\lambda\)'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The
values of \(\lambda\) describe how quickly the abundance of large body
sizes decline within a community. For example, a value of -1.5 means
there would be a relatively high number of large body sizes (shallow)
whereas a value of -2.5 means there would be relatively very few large
body sizes (steep).

@@

After simulating data, we used three different methods (described below)
to estimate the value of \(\lambda\) (maximum likelihood, equal
logarithmic bins, and normalized abundance spectra). We repeated this
process 1000 times (reps) and plotted the distribution of values
obtained for each method.

\hypertarget{maximum-likelihood-estimation-mle}{%
\subsection{Maximum likelihood estimation
(MLE)}\label{maximum-likelihood-estimation-mle}}

We estimated the exponent \(\lambda\) directly using MLE methods
modified from the the \texttt{sizeSpectra} package (Edwards et al.
2017). Throughout the manuscript, these estimates are referred to as
MLE.

@@

In addition, to estimate the OLS slope parameter in log-log space, we
used two common binning methods. . It is important to note that the
log-log regression is estimating \(\lambda\) as the regressionslope
(\(\beta_{1}\)), and not \(\lambda\) directly. In the interest of
clarity, we will refer to all estimates of lambda as \(\lambda\)
regardless of the method used, and reserve the terms ``parameters,''
``intercept,'' and ``slope'' \(\beta_{env}\) to refer to the parameters
of linear regression that uses the lambdas as a response variable.

For the first binning method, we created 6 equal logarithmic bins
covering the range of body sizes. This method has been used extensively
in previous studies. For example see (Dossena et al. 2012, Martínez et
al. 2016, Perkins et al. 2018). In the present study, the count in each
bin was normalized by dividing by the bin width to account for the
unequal bin sizes. The process of normalizing shifts the estimate by -1.
In other words, an un-normalized estimate of -0.75 would result in an
estimate of -1.75 when normalizing the data (Sprules and Barth 2015,
Edwards et al. 2017, Pomeranz et al. 2022). Throughout the manuscript,
the normalized equal logarithmic binning method will be referred to as
ELBn.

@@

The second method was similar to ELBn, but bins of log\textsubscript{2}
widths are used. The number of bins is dependent on the range of body
sizes present in the data. When working with empirical data with
different size ranges this can alter the number of bins per site, and it
is recommended to construct the bins based on the global min and max
body size present. However, since the data here were simulated from a
known size range, the number of bins for each site is identical. The
count in each bin is normalized in the same way as described above. This
method has been used extensively in the literature e.g., (Jennings et
al. 2002, Jennings and Blanchard 2004, Sprules and Barth 2015, McGarvey
and Kirk 2018, Pomeranz et al. 2019a) and is referred to as the
Normalized Abundance Spectrum (NAS).

For the two binning methods, simple OLS regression were conducted in the
form \[Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)\]

Where \(N\) is the count in each bin, \(M\) is the mid-point of each
bin, and \(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).

@@

@@

To do this, we first simulated \(\lambda\)'s from a linear regression
with a known slope \(\beta_{1}\) of -0.5 and intercept of -1.5 across a
generic predictor variable x with values ranging from -1 to 1:

\[\lambda_{ik} = \beta_{0} + \beta_{0}x_{ik}\]

This produced \(k\) = 1000 values from each value of \(x\) (-1, -0.5, 0,
0.5, 1), resulting in 5000 total \(\lambda\)'s. From each \(\lambda\) we
simulated 1000 individual body sizes using the procedure described
above.

Once the body size data were simulated, we then reversed the procedure.
First, we used the three methods (NAS, ELBn, MLE) to estimate
\(\lambda\) for each of the 5000 body size datasets. Then we fit a
linear regression to those data with the predictor variable \(x\) and
compared the resulting intercept (\(\beta_{0}\)) and slope
(\(\beta_{1}\)) parameter estimates to the known value. The main results
presented here were not dependent on the range of x-values or the number
of sites (Supplemental Information).

@@

We re-analyzed two data sets of benthic macroinvertebrate communities
from stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage (AMD) stress gradient. Details of the
sample collection and processing can be found in (Pomeranz et al.
2019a). Briefly, all individuals from each sample were identified to the
lowest practical taxonomic unit and body lengths were measured using
image processing software from photos taken with a camera mounted to a
dissecting microscope. Body mass was estimated using taxon-specific
published length weight regressions.

The second data set was from the wadeable stream sites of NEON (National
Ecological Observatory Network (NEON) 2022). NEON stream sites are
located across a wide temperature gradient in the United States, from
Puerto Rico to Alaska. Quantitative macroinvertebrate samples were
collected using the most appropriate method based on the local habitat.
All individuals were identified and had their body lengths measured, and
body mass was estimated using published length weight regressions. This
data has been analyzed previously using size spectra methods as
described in (Pomeranz et al. 2022). Detailed methods of the sampling
collection and data processing methods can be in the macroinvertebrate
DPI pubs found on the NEON website
\url{https://www.neonscience.org/data-collection/macroinvertebrates}.
Coefficient estimates (\(\pm\) 1 standard deviation) are compared across
methods. This allows us to determine whether or not the main results
would differ depending on the method used.

\hypertarget{performance-metrics}{%
\subsection{Performance Metrics}\label{performance-metrics}}

We compared performance of each procedure (NAS, ELBn, MLE) by first
plotting the distribution of lambda or regression parameter estimates
(\(\lambda\), \(\beta_{0}\), \(\beta_{1}\)) against the known values. We
calculated bias for the procedures overall as the median absolute
difference (averaged across all simulations) between the known values
and the modeled estimates. Finally, for each procedure we estimated the
width of the 95\% CI's to compare uncertainty.

@@

The data used in this manuscript is already publicly available (Pomeranz
et al. 2019b) data dryad DOI
\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s},
(National Ecological Observatory Network (NEON) 2022). R Scripts to
reproduce the full simulation and analysis are available at:
({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).
(To be archived upon acceptance.)

@@

There was considerable variation in the \(\lambda\) estimate across
methodologies (Figure 1). The distribution of estimates from the MLE
method was always symmetrical and centered at the known value of
\(\lambda\) (Figure 1). The distribution of estimates from the binning
methods were generally wider and occasionally asymmetrical or bimodal.
On average, the 95\% confidence intervals produced by the NAS and ELBn
methods were \textasciitilde2 times wider than those produced by MLE
(Table 1), indicating greater consistency of estimates from MLE.
Similarly, estimates of \(\lambda\) deviated from the true \(\lambda\)
by an average of 0.04 or 0.05 absolute units for the NAS and ELBn
methods, more than 4 times higher than deviation (0.01) for the MLE
(Table 1).

@@

\begin{longtable}[]{@{}llrrrr@{}}
\caption{Table 1. Summary of three methods in recapturing the known
values of lambda or the regression slopes simulated in this study.
Performance is determined by comparinged by uncertainty (range of 95\%
CI's) and the absolute distance of the model estimates from the known
estimates (median and sd of the difference). Values are summarized
across all n = 9000 or 6000 simulated datasets. See figures for more
specific comparisons.}\tabularnewline
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Target\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
n\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
Median Range of 95\% CI\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
Median Absolute Deviation\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
SD Absolute Deviation\strut
\endfirsthead
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Target\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
n\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
Median Range of 95\% CI\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
Median Absolute Deviation\strut
\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
SD Absolute Deviation\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Lambda\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
9000\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
0.0661400\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
0.0120454\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
0.0191500\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Lambda\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
NAS\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
9000\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
0.1335267\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
0.0460093\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
0.0929513\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Lambda\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
ELBn\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
8579\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
0.1672598\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
0.0350186\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
0.0675122\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Regression Slope\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
6000\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
0.0382185\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
0.0086982\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
0.0089766\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Regression Slope\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
NAS\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
6000\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
0.0799201\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
0.0513434\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
0.0442020\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
Regression Slope\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
ELBn\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
6000\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
0.0923495\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
0.0300186\strut
\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
0.0353218\strut

@@

The MLE method (Figure 3, blue) recaptured the known slope value in each
of the ``windows.'' with median absolute deviation of
\textasciitilde0.01 units (Table 1). By contrast, the binning methods
systematically overestimated the known slope (Figure 3), with median
absolute deviations \textgreater4 times higher than the MLE. Similarly,
uncertainty in the slope estimates derived from binning methods was
twice that of the uncertainty in the MLE method (Table 1).

@@

All methods recaptured the correct sign of the slopes, yielding
qualitative consistency (Figure 4). However, the binning methods again
systematically underestimated the true value of the slope by
\textasciitilde0.05 units (Figure 5). Likewise, uncertainty in the slope
estimates was always greater in the binning methods, with the width of
the distributions increases with stronger relationships across a
hypothetical gradient. By comparison, the MLE showed no evidence of bias
and was always centered at the known value with relatively narrow
variation.

\begin{figure}
\centering
\includegraphics{figures/vary_beta_plot.png}
\caption{Figure 4. Individual regressions for each rep (N = 1000) and
method (columns) for three different known relationship values (rows
from bottom to top): 0, 0.25, and 0.5}
\end{figure}
